---
title: "STAC33-TUT5"
author: "Uzair Mirza"
date: "13/02/2022"
output: pdf_document
---
# Introduction
This week's lecture we will be discussing topics from CH:9, Ch:10 and Ch:12. Chapter is about **Sign test** and Chapter 10 is about **Mood Median test** and Chapter 12 is about **QQ-plot**(aka normal quantile plots). All the problems being discussed can be found on the PASIAS [here](https://ritsokiguess.site/pasias/)

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
```

## Question 9.10
Data is about amount of protein in the package. The advertisement claims to have **6 ounces/pack**(our assumption about the package $H_{0}$). Goal is to test this claim by sampling(n=20) and conducting some tests.

### a + b. Read the data and make sutiable plot 
```{r, message=FALSE}
# a
my_url <- "http://ritsokiguess.site/datafiles/protein.txt"
## Data is not seperated by a space/chracter or tab so this works
meals <- read_table(my_url)

# b
## 1 quantitative variable(Can make box-plot/histogram)
## Interested in distribution rather than range so histogram
ggplot(meals, aes(x=protein)) + geom_histogram(bins = 5)
```
Note we can observe? \
- (Not symmetric) \
- Skewed left \

### c. Why might a sign test be better than a  t-test for assessing the average amount of protein per package? Explain briefly. (“Average” here means any measure of centre.) 
Assumptions for t-test: \
- Normality \
Here we can see the assumption is violated(as not normal from Histogram). \
- We testing $\mu$ in t-test and here it is not a reliable measure(think about the sample not being accurate and $\mu$) as it is sensitive. \
- Sample size not good enough for LLN(n=20) here \

Hence a sign test on Median is a better test to use. \
Sign test has no dependencies/assumptions about the population parameters or the distributions \
Can perform test on Median. \
Median, Mode not sensitive to odd values(outliers). \

### d. Run a suitable sign test for these data. What do you conclude? 
**Setup, !!!THIS IS WHAT GOOGLE WOULD SUGGEST, DON'T DO THIS!!!** \
```{r, eval=FALSE}
install.packages('smmR')
```

**Follow this from lecture slides.** \
github source: [here](https://github.com/nxskok/smmr) \
```{r, eval=FALSE}
# One time setup
install.packages('devtools')
library(usethis) # this step is specific to my machine, might not be required for you
library(devtools)
install_github("nxskok/smmr")
```

```{r, message=FALSE}
library(smmr)
```
This is how you run the test. \
Now note we are interested in the the Median being = 6, so we state that. Furthermore think about whether it will be 2-sided or 1 sided? \
- 2-sided bc the value can lie on either side. \
`sign_test(data, variable, $H_{0}$)` $\alpha = 0.05$
```{r}
sign_test(meals, protein, 6)
```
The P-value, 0.0414, is less than $\alpha = 0.05$ 0.05, so we reject the null hypothesis and conclude that the median is different from 6 ounces. The advertisement by the company is not accurate. \

### e. In your sign test, how could you have deduced that the P-value was going to be small even without looking at any of the P-values themselves?
- How is Sign test conducted ? \
  - Compare the sampled values against the $H_{0}$ \
    - What happens when values are EQUAL? \
    Don't include those values as currently interested in `>` or `<` 
```{r}
meals %>% 
  group_by(protein > 6) %>% 
  summarise(n = n())
```
  - Note 15 are less than 6. Hence can reject and assume p-value will be less without conducting the test. \

### f. Obtain a 90% confidence interval for the population median protein content. What does this tell you about the reason for the rejection or non-rejection of the null hypothesis above?
`ci_median(data, variable,conf.level = $1-\alpha$ )` by default $\alpha = 0.05$
```{r}
alpha = .1
ci_median(meals, protein, conf.level = 1-alpha)
```
- Does our $H_{0}$ value lie withing the range at $\alpha = 0.1$? \
  - No! 6 not $\in$ CI. \
  
## 10.8 Handspans revisited 
Recall the original study was to compare the handspan between males and females. \
So we have 2 groups in our study. \

### a + b. Load data make plots to investigate and compare normality between the sample group
```{r}
# data
my_url <- "http://ritsokiguess.site/datafiles/handspan.txt"
span <- read_delim(my_url, " ")

# QQ-Plot
ggplot(span, aes(sample=handspan)) + stat_qq() + stat_qq_line() + facet_wrap(~sex)
```
- What is QQ plot? \
  - What is a quantile ?
    Can think **similar**(but not really) to bins. \
  - How is it plotted? \
    Compare the quantiles of Normal(0,1) against the sample's quantiles \
- Our plot description: \
  - Males: \
    Not normal as the values to the left are deviating away from the normal line(SLIGHTLY left skewed) \
  - Females: \
    Same as males + seem to have outliers aswell on the left side \
    

### c. Discuss briefly whether you might prefer to use Mood’s median test to compare the handspans of the male and female students, compared to a two-sample t-test.
- T-test assumptions about the normality of the sample is not satisfied. \
- Mood-Median is the move as does not rely on $\mu$ + Normality. \

### d. Run the Mood-Median test
$$H_{0}: \mu_{\frac{1}{2}} = 0 \qquad \Longrightarrow \mu_{\frac{1}{2}} = median $$

The median between both the groups is the same ie the difference is 0. \
**Setup** ***the call for installing the package is covered above.***
```{r, eval=FALSE}
# library call
library(smmr)
```

To run the test. `median_test(data, variable of the median, group)` \
```{r}
median_test(span, handspan, sex)
```
The P-value of `2.66404e-19	`is extremely small, so we can conclude that males and females have different median handspans. Remember that we are now comparing medians, and that this test is **two-sided**. \

## 12.3 Lengths of heliconia flowers
Flower length and the beak of the humming-bird has been evolved according to the specie

### a,b,c,d. Read data and make QQ-Plot for each specie
```{r}
# a
my_url <- "http://ritsokiguess.site/datafiles/heliconia.csv"
heliconia = read_csv(my_url)

# bcd
## way 1
ggplot(heliconia,aes(sample=caribaea_red)) + stat_qq() + stat_qq_line()

## way 2
heliconia %>% ggplot(aes(sample=bihai)) + stat_qq() + stat_qq_line()

ggplot(heliconia,aes(sample=caribaea_yellow))+stat_qq()+stat_qq_line()

```
### e, f. Which one is the clossest to the normal and of the other explain why they are not.
- `caribaea_yellow` closest to normal as the quantiles seems to match with Normal QQ.
- `bihai` seems to have a S-shaped if we negate the outliers at the top + left skewed with removal of outliers and without the removal seems to be right skewed.
- `caribaea_red`  
    Short tailed
    Initially points seem too bunched up on the line then they seem to tail away from the QQ-line. Looks Uniform distribution. \

```{r}
### Source: https://bioinfo.iric.ca/permutations/

x <- rnorm(1000, mean=50, sd=5) # normal distribution
z <- runif(1000) # uniform distribution 
par(mfrow=c(2,2))
qqnorm(x, pch=20)
qqline(x)
hist(x)
qqnorm(z, pch=20)
qqline(z)
hist(z)

```

Check more examples of QQ x Distribution relation. [here](https://towardsdatascience.com/q-q-plots-explained-5aa8495426c0)