---
title: "TUT-5 Live"
author: "Uzair"
date: "2/13/2022"
output: pdf_document
---
# Introduction
This week's lecture we will be discussing topics from CH:9, Ch:10 and Ch:12. Chapter is about **Sign test** and Chapter 10 is about **Mood Median test** and Chapter 12 is about **QQ-plot**(aka normal quantile plots). All the problems being discussed can be found on the PASIAS [here](https://ritsokiguess.site/pasias/)

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
```

## Question 9.10
Data is about amount of protein in the package. The advertisement claims to have **6 ounces/pack**(our assumption about the package $H_{0}$). Goal is to test this claim by sampling(n=20) and conducting some tests.

### a + b. Read the data and make sutiable plot 
```{r, message=FALSE}
# a
my_url <- "http://ritsokiguess.site/datafiles/protein.txt"
## Data is not seperated by a space/chracter or tab so this works
meals <- read_table(my_url)
head(meals)

# b
## 1 quantitative variable(Can make box-plot/histogram)
## Interested in distribution rather than range or outliers so histogram
```
Note we can observe? \
- 

### c. Why might a sign test be better than a  t-test for assessing the average amount of protein per package? Explain briefly. (“Average” here means any measure of centre.) 
Assumptions for t-test: \


### d. Run a suitable sign test for these data. What do you conclude? 
**Setup, !!!THIS IS WHAT GOOGLE WOULD SUGGEST, DON'T DO THIS!!!** \
```{r, eval=FALSE}
install.packages('smmR')
```

**Follow this from lecture slides.** \
github package source: [here](https://github.com/nxskok/smmr) \
```{r, eval=FALSE}
# One time setup
install.packages('devtools')
library(usethis) # this step is specific to my machine, might not be required for you
library(devtools)
install_github("nxskok/smmr")
```

```{r, message=FALSE}
library(smmr)
```
This is how you run the test. \

`sign_test(data, variable, $H_{0}$)` $\alpha = 0.05$
```{r}
```


### e. In your sign test, how could you have deduced that the P-value was going to be small even without looking at any of the P-values themselves?
- How is Sign test conducted ? \


```{r}

```

### f. Obtain a 90% confidence interval for the population median protein content. What does this tell you about the reason for the rejection or non-rejection of the null hypothesis above?

```{r}

```
- Does our $H_{0}$ value lie withing the range at $\alpha = 0.1$? \

  
## 10.8 Handspans revisited 
Recall the original study was to compare the handspan between males and females. \
So we have 2 groups in our study. \

### a + b. Load data make plots to investigate and compare normality between the sample group
```{r}
# data
my_url <- "http://ritsokiguess.site/datafiles/handspan.txt"
span <- read_delim(my_url, " ")

# QQ-Plot

```
- What is QQ plot? \
  - What is a quantile ?
  
  - How is it plotted? \
  
- Our plot description: \
  - Males: \
   
  - Females: \
    

### c. Discuss briefly whether you might prefer to use Mood’s median test to compare the handspans of the male and female students, compared to a two-sample t-test.
- T-test assumptions about the normality of the sample is ___ satisfied. \

### d. Run the Mood-Median test
$$H_{0}: \mu_{\frac{1}{2}} = 0 \qquad \Longrightarrow \mu_{\frac{1}{2}} = median $$

The median between both the groups is the same ie the difference is 0. \
**Setup** ***the call for installing the package is covered above.***
```{r, eval=FALSE}
# library call
library(smmr)
```

To run the test. `median_test(data, variable of the median, group)` \
```{r}
```


## 12.3 Lengths of heliconia flowers
Flower length and the beak of the humming-bird has been evolved according to the specie

### a,b,c,d. Read data and make QQ-Plot for each specie
```{r}
# a
my_url <- "http://ritsokiguess.site/datafiles/heliconia.csv"
heliconia = read_csv(my_url)

# bcd
## way 1

## way 2

```
### e, f. Which one is the clossest to the normal and of the other explain why they are not.
- `caribaea_yellow` 
- `bihai` 
- `caribaea_red`  
    
    


```{r}
### Source: https://bioinfo.iric.ca/permutations/

x <- rnorm(1000, mean=50, sd=5) # normal distribution
z <- runif(1000) # uniform distribution 
par(mfrow=c(2,2))
qqnorm(x, pch=20)
qqline(x)
hist(x)
qqnorm(z, pch=20)
qqline(z)
hist(z)

```

Check more examples of QQ x Distribution relation. [here](https://towardsdatascience.com/q-q-plots-explained-5aa8495426c0)