---
title: "Live TUT-2"
author: "Uzair"
date: "1/24/2022"
output: html_document
---

# Introduction


The chapters can be found here: [https://ritsokiguess.site/pasias/drawing-graphs.html] \
Notes to follow along are here: [https://github.com/ujji7/STAC33-TUT] \


## CH:3.6 Juice problem
First load the library tidyverse
```{r, load_libraries, include=FALSE, results='hide'}
library(tidyverse)
```


### Load data + look at some of the values
url <- "http://ritsokiguess.site/datafiles/ojuice.txt"
```{r}

```



### Goal: find a relation between the 2 variables
```{r}

```



### Summary


## 3.7 Making soap
```{r, warning=FALSE}
url <- "http://ritsokiguess.site/datafiles/soap.txt"
soap <- read_delim(url, " ")
```

Lets look at some of the data now along with some properties of our data
`head()` -> 
`summary()` ->
```{r, warning = F}

```


### Make a histogram for SCRAP
To make histogram we use `geom_histogram()` with an appropriate binsize \
```{r}

```



3) Features of the graph
  **modes**
  **Range** 
  **Skewness?** 

## Make a side-by-side box-plot
`geom_boxplot()` is used to make the box-plot
```{r}

```





3) Features of the graph
  **Range** between the 2 plants \
  **Median**, \
  **outlier** \
  **MEAN** 
  
  
  
  
  
### Summary

## 4.11

### Reading the data and observing some of the data
```{r}
my_url <- "https://raw.githubusercontent.com/nxskok/datafiles/master/compatt.txt"
anxiety <- read_delim(my_url," ")

head(anxiety)
```

### Number of males and females in our sample
```{r}

```















`group_by()` 
`summarise()` ->  `n()` 

### Mean and SD of CAS and CARS not seperated by gender
```{r}

```



Doing it without mentioning the column name
```{r}

```


`mean()`, `sd()` -> 
`select()` -> 
`across()` -> 
`list()` ->

Same thing for all the quantitative variables
```{r}

```



`where()` -> 
`is.numeric`-> 

Again to see and compare the relation between the 2 variables in each group
```{r}

```


`color()` -> 


### Add and store the sum of CARS and CAS and compare the average of this SUM now between Male and Female
```{r, warning=F}
```


`mutate()` -> 

### Summary
`group_by()` -> 
`summarise()` -> 
`where()` -> 
