---
title: "TUT 7"
author: "Uzair Mirza"
date: "07/03/2022"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
```

# Introduction
This week's lecture we will be discussing topics from Ch:17. Chapter 17 is about **Tidying Data**. All the problems being discussed can be found on the PASIAS [here](https://ritsokiguess.site/pasias/)


## Q.17.22 Feeling the heat
There were heat alerts in Toronto and the city decided to open the pools and cooling stuff for extra hours. \
The dataset we have has all the heat alerts from 2001 to 2016. \

### a. Read the data and observe the variables
```{r}
my_url <- "http://ritsokiguess.site/datafiles/heat.csv"
heat <- read_csv(my_url)

view(heat)
```
Specifically in this dataset the new types of data is `date`. \
Note the data file contains the dates in **year-month-day** order hence read_csv is able to recognise this. \
Having dates in `date` format allows us to work with them in an easier way(we can  specify ranges and etc.) \

(if the source file did not follow the same format(**year-month-day**) `lubridate` could be used to read in the dates as dates) \

### b. Get the count for each `code`
```{r}
heat %>% group_by(code) %>% summarise(n = n()) -> a
a
```

What each code means:
```{r}
code <- a$code
meaning <- c("(Start of) Extended Heat Alert", "Extreme Heat Alert downgraded to Heat Alert", "Extended Heat Alert continues", "(Start of) Heat Alert", "Heat Alert continues", "Heat Alert upgraded to Extended Heat Alert")

codeXmeaning <-tibble(code, meaning)
codeXmeaning
```

### c. How many (regular and extended) heat alert events are there altogether? A heat alert event is a stretch of consecutive days, on all of which there is a heat alert or extended heat alert
Essentially we are trying to find when there was a change in `code` status for our recorded dates ie Event starts specefic generated code is issued, event ends the specific generated code is generated. \
So we are trying to find the case when the difference between these code changes was more than 1. \

```{r}
heat %>%
  select(-text) %>% #drops the variable
  mutate(daycount = as.numeric(date)) %>% # gets the date diffrence from a standard date(1970)
  mutate(daydiff = abs(c(diff(daycount), 0))) %>% #
  count(daydiff != 1) #
```

`as.numeric(date)` -> counts the number of days it has been since a specific date(some date in 1970) \
`diff()` -> takes a difference st; given tupple x our result y is generated such as. \ y[i] = x[i+1] - x[i]. \
For the last value we compare it with itself, hence diff is always 0.

### d. We are going to investigate how many heat alert days there were in each year.
To do that, we have to extract the year from each of our dates and work with the data \

Library `lubridate` is used to work with the dates here. \
Function `year()` extracts the year from the date and stores it an  `dbl` \

```{r}
# the library needed for the function call year()
library(lubridate)

# 
heat %>% select(-text) %>% mutate(year = year(date))
# gets the count of unique years
heat %>% select(-text) %>% mutate(year = year(date)) %>% count(year)
```


## 17.15 Ethanol and sleep time in rats
Effect of Ethanol on sleep time in rats. So we have 20 rats, 5 for each group of ethanol; we have data for 3 different ethanol levels and 1 control group.

### a. Read and observe the format of the data
```{r}
my_url <- "http://ritsokiguess.site/datafiles/ratsleep.txt"
sleep1 <- read_delim(my_url, " ")

view(sleep1)

```

So here each observation in the table has the value of the sleep time of the mice. \
Note it is not really useful in terms of testing when working with data in this format. Instead let us try to improve on it to get in in a format to carry out tests and make plots.

### b. Transform this data frame into one that you could use for modelling or making graphs.
GOAL: get a col so we have the sleep time for each observation under a specific variable `sleeptime` . \
`pivot_longer()` is the function to use. `it takes in a WIDE data frame and makes it LONGER; reduces the number of col and increases the number of rows

```{r}
(sleep1 %>% 
  pivot_longer(!treatment, names_to="rep", values_to="sleeptime") -> sleep)

(sleep1 %>% 
  pivot_longer(-treatment, names_to="rep", values_to="sleeptime") -> sleep)
```
`names_to= ` -> the col being concatenated into the new variable \
`value_to=` -> assigning to the new variable. \

### c. Using your new data frame, make side-by-side boxplots of sleep time by treatment group.
```{r}
ggplot(sleep, aes(x = treatment, y = sleeptime)) + geom_boxplot()
```

### d. In your boxplots, how does the median sleep time appear to depend on treatment group?
Differs between the groups... \


## 17.18 Location, species and disease in plants
Reading the table and what the variables are.\
2 Species, A and B \
Plausible location for the disease X and Y \
Disease present or absent  P and A. \
```{r}
my_url <- "http://ritsokiguess.site/datafiles/disease.txt"
tbl <- read_table(my_url)

view(tbl)
```

Lets make it tidy with the goal to make sense of the data. \
GOAL: We want to observe the frequencies in each category. So lets start with creating that variable. \
```{r}
tbl %>% pivot_longer(-Species, names_to="disloc", values_to = "frequency") -> tbl.2
tbl.2
```
Lets improve on it. \
GOAL: split `disloc` into 2 variables which can account for whether the disease is present or absent and the location of the disease. \
We notice it is a character of len 2, hence separation based on the first character will do the job. \
`seprate(variable, c("new_a","new_b"), c)`. Here the `variable` is the variable in the df we are splitting, `c("new_a","new_b")` are the new variables it is being split into. `c` is the length of the spiting point for the variable. 
```{r}
(tbl.2 %>% separate(disloc, c("disease", "location"), 1) -> tbl.3)
```



## More
So you can read about some more useful functions on the [website](https://ritsokiguess.site/pasias/tidying-data.html). \
`pivot_wider` -> Opposite is pivot longer, 17.19 Mating songs in crickets \
`unite` -> Joins values of col into a single col, 17.19 Mating songs in crickets \