---
title: "STAC33 TUT-2"
output:
  pdf_document: default
  html_notebook: default
---

# Introduction
We will be discussing problems from ch:3 and ch:4 from the PASIAS(Problems and Solutions in Applied Statistics). Ch:3 focuses on drawing graphs and ch:4 is about exploring data. So hopefully by the end of this session we are familiar with what graphs to draw given a problem and also perform introductory data analysis given a dataset.

The chapters can be found here: [https://ritsokiguess.site/pasias/drawing-graphs.html]

## CH:3.6 Juice problem
First load the library tidyverse
```{r, load_libraries, include=FALSE, results='hide'}
library(tidyverse)
```


### Load data
1) Specify the URL (optional BUT recomended) \
2) load data
  2.1) `read_delim()` -> used to load data separated by a certain character which needs to be specefied


```{r}
url <- "http://ritsokiguess.site/datafiles/ojuice.txt"
juice <- read_delim(url, " ")
```
Lets look at some of data
`glimpse()` -> shows the values along with characteristics
`head()` -> shows first few values
```{r}
glimpse(juice)
head(juice)
```
 - Variables are -> **run, sweetness, pectin**

### Goal: find a relation between the 2 variables
1) Note the TYPE of the variable involved. (Quantitative, sweetness appears to be continuous)
2) Choose Scatter plot to observe a trend

3) Then Draw a line between the plot to see the relation

`ggpplot()` -> plotting package in tidyverse that provides helpful commands to create plots
Think of this as a masterchef you call for a day who will be helping you make your meal, this chef will setup the place and the setting for you but you have to make the choice of using a knife vs a peeler depending on the task
```{r}
# just the scatter plot
ggplot(juice, aes(x = pectin, y = sweetness)) + geom_point()

# with the line of best-fit
ggplot(juice, aes(x = pectin, y = sweetness)) + geom_point() + geom_smooth(method = lm, se= F)
## method = lm for best fit
## se = F removes the confidence bands around the line
```
- General downwards trend with the increase in pectin
- But not really convincing as there appears to be 2 extreme values above 350 BUT they are also follow the same trend

### Summary
- When asked to come up with the relationship between 2 quantities(both or atleast 1 being Quantitative) we make a scatter plot and observe the trend between them

## 3.7 Making soap
```{r, warning=FALSE}
url <- "http://ritsokiguess.site/datafiles/soap.txt"
soap <- read_delim(url, " ")
```

Lets look at some of the data now along with some properties of our data
`head()` -> shows first few values
`summary()` -> characteristics of the data
`table()` -> shows the different UNIQUE values along with their counts, under a variable preferbly used for CATAGORICAL variables
```{r, warning = F}
head(soap)
summary(soap)

# I use this as I found it on stackoverflow once don't use this when asked to 
# show the result as the generated output is not clean 
table(soap$line)
```


### Make a histogram from SCRAP
To make histogram we use `geom_histogram()` with an appropriate binsize \

1) what are bins? \
The count you specify for the bins is the basic partition you are willing to make for your x values \
2) Ideal number? \
theory vs application. just play around and fit what is ideal \
make sure it's never more than your actual sample size :p

```{r}
soap %>% ggplot(aes( x= scrap)) + geom_histogram(bins = 10)
```
3) Features of the graph
  **Bimodail** ? type of modes
  **Range** 
  **Skewness?** Sym

## Make a side-by-side box-plot
`geom_boxplot()` is used to make the box-plot
```{r}
ggplot(soap, aes(x = line , y = scrap)) + geom_boxplot()
```
3) Features of the graph
  **Range** between the 2 plants \
  **Median**, range of a > b \
  No **outlier** \
  BUT there is a huge overlap between their ranges \
  Never make inference about **MEAN** when discussing box-plot NOT reliable \
  
### Summary
Histogram for distribution and spread and mode \
bins are for dividing the range of x-values \
Box-plot for range if values + Q1 Q3 median + outliers \
  

## 4.11

### Reading the data and observing some of the data
```{r}
my_url <- "https://raw.githubusercontent.com/nxskok/datafiles/master/compatt.txt"
anxiety <- read_delim(my_url," ")

head(anxiety)
```

### Number of males and females in our sample
```{r}
# prefered way to use RESULT is a tibble
anxiety %>% group_by(gender) %>% summarise(count = n()) 

# my hacky way from stackoverflow RESULT is a console output NOT RECOMENDED for displaying results
table(anxiety$gender) 
```
`group_by()` -> separates our data into different groups based on the specified variable
`summarise()` -> used to make meaning summaries by declaring a new variable here we used `n()` and named that variable count

### Mean and SD of CAS and CARS not seperated by gender
```{r}
anxiety %>% summarise(mean_CAS = mean(CAS), mean_CARS = mean(CARS), sd_CAS = sd(CAS), sd_CARS = sd(CARS))
```
Doing it without mentioning the column name
```{r}
anxiety %>% select(2,3) %>%
  summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))))
```
`mean()`, `sd()` -> calculates the mean and sd
`select()` -> to choose the column with the position of the column in the dataframe
`across()` -> to go over what the selected variables have in common
`list()` -> used because we are doing multiple opoerations on each variable

Same thing for all the quantitative variables
```{r}
anxiety %>% summarize(across(where(is.numeric), list(m = ~mean(.), s = ~sd(.))))
```
`where()` -> used to declare a conditional statement
`is.numeric` -> the variable TYPE is numeric

Again to see and compare the relation between the 2 variables in each group
```{r}
#general trend
ggplot(anxiety,aes(x=CAS,y=CARS)) + geom_point() + geom_smooth(method = lm, se=F)
# trend in each gender group
ggplot(anxiety,aes(x=CAS,y=CARS, color = gender)) + geom_point() + geom_smooth(method = lm, se=F)
```
`color()` -> to separate the types in the categorical variable


### Add and store the sum of CARS and CAS and compare the average of this SUM now between Male and Female
```{r, warning=F}
anxiety %>% mutate(sum = CARS + CAS) %>% group_by(gender) %>% summarise(avg = mean(sum))
```
with `mutate()` we are first creating this new variable and then with the `summarise()` we are finding the mean of the sum for each gender.\
Here note we are not storing any of these changes(notice environment) right now we are just temporarily working with them and discarding them after processing our results \
To actually store the changes just declare it to a new variable. \

`mutate()` -> used to make a new variable. You declare the name of the new variable and the assignment rule for that variable

### Summary
`group_by()` -> use on catagorical variables and think of it and typing down knots based on the unique groups \
`summarise()` -> creates summaries like mean mode median ... \
`where()` -> used for conditional statements think \
