---
title: "Live TUT-3"
author: "Uzair"
date: "1/31/2022"
output: html_document
---

# Introduction
We will be discussing problems from CH:6 and CH:7 from [PASIS](https://ritsokiguess.site/pasias/one-sample-inference.html) during this week's TUT. \
CH:6 focuses on 1-sample test while CH:7 focuses on 2-sample test. \

## Example 6.7 Hunter-gatherers in Australia
### load tidyverse
```{r, echo=F, include=FALSE}
library(tidyverse)
```

### a. Load Data and see how many societies were studied?
```{r}
source ="http://ritsokiguess.site/datafiles/hg.txt"

```
Now let's find the count of these unique societies \
```{r}

```

### b. Test whether these Australian hunter-gatherer societies are like the rest of the world in terms of mean population density. State suitable null and alternative hypotheses. \

FACT:  Anthropologists have studied hunter-gatherer societies in forest ecosystems across the world. The average population density of these societies is **7.38** people per $100 km^2$ \

Hypothesis: \
**NULL HYPOTHESIS** \
$$H_{0}: \mu  $$
**ALTERNATIVE HYPOTHESIS**
$$H_{a}: \mu $$


### c. Test your hypotheses using a suitable test. What do you conclude, in the context of the data?

```{r}

```
  - p-value = 
    - 
  - 

### d. Validity of the results?
Assumptions for t-test:
```{r, plots}

```
 
 
 
```{r, qq}

```
  
  

### e. Case when $H_{a} < c$
Suppose that:
$$H_{a} < 7.38$$
In this case we will be using a 1-sided t-test 
```{r}

```
  

## 7.13 Handspans of males and females
### a. Load and observe the data and make the a sutaible plot to display the range for handspan for each group.
```{r}
my_url <- "http://ritsokiguess.site/datafiles/handspan.txt"

# view some of the data

# make the suitable plot

```
 

### b. Test the hypothesis that M $\geq$ F
$$H_{a} = \mu_{M} < \mu_{F} \Longrightarrow \mu_{M} - \mu_{F} < 0 $$

$$ \mu_{a} = \mu_{M} - \mu_{F} \Longrightarrow H_{a} = \mu_{a} < 0 $$
Need to test:
$$ H_{0} \geq 0 \quad    \quad H_{a} < 0 $$
```{r, test_2}

```


### c. Consutruct a confidence interval(CI) at $\alpha = 0.1$
What is CI? What is this $\alpha = 0.05$ \


So now for this example, lets do the test:
```{r}


```
On a question like this, the grader is looking for three things:
  - getting the output
  - saying what the interval is
  - rounding it to a suitable number of decimals.



## Common issues
  - START EARLY!!
  - Won't knit: Try cloud version, code-chuck issue -> eval = False
  - loading data-set -> check out diff ways to load data
  - IF dataset is super big then never call the data-set in a codechuck(unless asked for it) it will take super long to knit + submission will be super long
  - ONLY state the assumptions you're sure about WITH proper reasoning.